### 2017-02-14 docs/devel.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  devel.tex
%
%  Information for users and delvelopers
%
%  Edinburgh Soft Matter and Statistical Physics Group and
%  Edinburgh Parallel Computing Centre
%
%  (c) 2014 The University of Edinburgh
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  devel.tex
%
%  Information for users and delvelopers
%
%  Edinburgh Soft Matter and Statistical Physics Group and
%  Edinburgh Parallel Computing Centre
%
%  (c) 2014-2017 The University of Edinburgh
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsubsection{Array address models}

In a related issue, the code allows different data layout to be used
to improve performance. The default layout is array of structures (AOS)
which is suitable for most CPU architectures. (A blocked version AOSOA
is also available which may promote vectorisation.) For GPU machines, a
structure of arrays (SOA) format can be selected at compile time.

@@

\subsubsection{Address models}

The default memory layout is array of structures (AOS). Compiler
preprocessor macros for blocked SOA (\texttt{-DADDR\_AOSOA}) or
structure of arrays (\texttt{-DADDR\_SOA})
may be appropriate for different platforms.

Add the relevant flags to the \texttt{config.mk} file before
compilation to effect the given address model.

### 2017-02-14 docs/develop.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  develop.tex
%
%  Further information for delvelopers
%
%  Edinburgh Soft Matter and Statistical Physics Group and
%  Edinburgh Parallel Computing Centre
%
%  (c) 2014-2015 The University of Edinburgh
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The code allows for different order of storage associated with 3-dimensional
fields (scalars, vectors, and so on). For historical reasons these are
referred to as `Model' and `Model~R' options, which correspond
to array-of-structures and structure-of-arrays layout, respectively. Early
versions of the code for CPU were written to favour summation of LB
distributions on a per lattice site basis in operations such as
$\rho(\mathbf{r}) = \sum_i f_i(\mathbf{r})$. This is array-of-structures,
where the $f_i$ are stored contiguously per site. Introduction of GPU
versions, where memory coalescing favours the opposite memory order,
were then referred to as `Model~R', the `R' standing for `reverse'.

@@

\textbf{ADDR\_MODEL}: The array-of-structures or Model order for an
$n-$vector field
with components $ v_\alpha = (v_0, v_1, \ldots, v_{n-1})$ is, schematically:
\[
\underbrace{ \boxed{v_0\vphantom{v_{n-1}}} \boxed{v_1\vphantom{v_{n-1}}}
\boxed{\ldots\vphantom{v_{n-1}}} \boxed{v_{n-1}} }_{r_0}
\underbrace{ \boxed{v_0\vphantom{v_{n-1}}} \boxed{v_1\vphantom{v_{n-1}}}
\boxed{\ldots\vphantom{v_{n-1}}} \boxed{v_{n-1}} }_{r_1}
\underbrace{ \boxed{v_0\vphantom{v_{n-1}}} \boxed{v_1\vphantom{v_{n-1}}}
\boxed{\ldots\vphantom{v_{n-1}}} \boxed{v_{n-1}} }_{r_2} \ldots
\underbrace{ \boxed{v_0\vphantom{v_{n-1}}} \boxed{v_1\vphantom{v_{n-1}}}
\boxed{\ldots\vphantom{v_{n-1}}} \boxed{v_{n-1}} }_{r_{n_s-1}}
\]

\textbf{ADDR\_MODEL\_R}: The Model~R version is:
\[
\overbrace{
\begin{array}{cc} \boxed{v_0} \\  _{r_0} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{v_0} \\  _{r_1} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{\ldots\vphantom{v_0}} \\ \phantom{r_2}\end{array} \mkern-22mu
\begin{array}{cc} \boxed{v_0} \\  _{r_{n_s-1}} \end{array}
}^{component_0}
\overbrace{
\begin{array}{cc} \boxed{v_1} \\  _{r_0} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{v_1} \\  _{r_1} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{\ldots\vphantom{v_0}} \\ \phantom{r_2}\end{array} \mkern-22mu
\begin{array}{cc} \boxed{v_1} \\  _{r_{n_s-1}} \end{array}
}^{component_1}
\begin{array}{cc} \ldots \\ \phantom{r_2} \end{array}
\overbrace{
\begin{array}{cc} \boxed{v_{n-1}} \\  _{r_0} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{v_{n-1}} \\  _{r_1} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{\ldots\vphantom{v_{n-1}}} \\ \phantom{r_2}\end{array} \mkern-17mu
\begin{array}{cc} \boxed{v_{n-1}} \\  _{r_{n_s-1}} \end{array}
}^{component_{n-1}}
\]

@@

\textbf{ADDR\_MODEL}:
A general rank 2 tensor $t_{\alpha\beta}$ with components $(t_{0,0},
\ldots, t_{m-1,n-1})$ is stored as:
\begin{gather*}
\underbrace{
\boxed{t_{0,0}\vphantom{t_{m-1,n-1}}} \boxed{t_{1,0}\vphantom{t_{m-1,n-1}}}
\boxed{\ldots\vphantom{t_{m-1,n-1}}} \boxed{t_{m-1,0}\vphantom{t_{m-1,n-1}}}
\boxed{t_{0,1}\vphantom{t_{m-1,n-1}}} \boxed{t_{1,1}\vphantom{t_{m-1,n-1}}}
\boxed{\ldots\vphantom{t_{m-1,n-1}}} \boxed{t_{m-1,1}\vphantom{t_{m-1,n-1}}}
\ldots
\boxed{t_{0,n-1}\vphantom{t_{m-1,n-1}}}\boxed{t_{1,n-1} \vphantom{t_{m-1,n-1}}}
\boxed{\ldots\vphantom{t_{m-1,n-1}}} \boxed{t_{m-1,n-1}\vphantom{t_{m-1,n-1}}}
}_{r_0}
\ldots
\\
\ldots
\underbrace{
\boxed{t_{0,0}\vphantom{t_{m-1,n-1}}} \boxed{t_{1,0}\vphantom{t_{m-1,n-1}}}
\boxed{\ldots\vphantom{t_{m-1,n-1}}} \boxed{t_{m-1,0}\vphantom{t_{m-1,n-1}}}
\boxed{t_{0,1}\vphantom{t_{m-1,n-1}}} \boxed{t_{1,1}\vphantom{t_{m-1,n-1}}}
\boxed{\ldots\vphantom{t_{m-1,n-1}}} \boxed{t_{m-1,1}\vphantom{t_{m-1,n-1}}}
\ldots
\boxed{t_{0,n-1}\vphantom{t_{m-1,n-1}}}\boxed{t_{1,n-1} \vphantom{t_{m-1,n-1}}}
\boxed{\ldots\vphantom{t_{m-1,n-1}}} \boxed{t_{m-1,n-1}\vphantom{t_{m-1,n-1}}}
}_{r_{n_s-1}}
\end{gather*}
Dyadic tensors, for example the gradient of a vector field
$\partial_\alpha v_\beta$ in three dimensions, are stored in corresponding
fashion with $m=3$ and
$\partial_\alpha = (\partial_x, \partial_y, \partial_z)$.

\textbf{ADDR\_MODEL\_R}: The Model R version is
\[
\overbrace{
\begin{array}{cc} \boxed{t_{0,0}} \\  _{r_0} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{t_{0,0}} \\  _{r_1} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{\ldots\vphantom{t_{0,0}}} \\ \phantom{r_2}\end{array} \mkern-18mu
\begin{array}{cc} \boxed{t_{0,0}} \\  _{r_{n_s-1}} \end{array}
}^{component_{0,0}}
\overbrace{
\begin{array}{cc} \boxed{t_{1,0}} \\  _{r_0} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{t_{1,0}} \\  _{r_1} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{\ldots\vphantom{t_{0,0}}} \\ \phantom{r_2}\end{array} \mkern-18mu
\begin{array}{cc} \boxed{t_{1,0}} \\  _{r_{n_s-1}} \end{array}
}^{component_{1,0}}
\begin{array}{cc} \ldots\vphantom{t_{n-1}} \\ \phantom{r_2}\end{array}
\overbrace{
\begin{array}{cc} \boxed{t_{m-1,n-1}} \\  _{r_0} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{t_{m-1,n-1}} \\  _{r_1} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{\ldots\vphantom{t_{0,0}}} \\ \phantom{r_2}\end{array} \mkern-17mu
\begin{array}{cc} \boxed{t_{m-1,n-1}} \\  _{r_{n_s-1}} \end{array}
}^{component_{m-1,n-1}}
\]

@@

The general rank 3 object $t_{\alpha\beta\gamma}$ with components
$(t_{0,0,0}, \ldots, t_{m-1,n-1,p-1})$ is stored in a manner which
generalises from the above, i.e., with the rightmost index running
fastest. Diagrams are omitted, but Model and Model R storage patterns
follow the same form as seen above.

@@

\textbf{ADDR\_MODEL}: For the LB distributions, up to two distinct
distributions can be accommodated to allow for a binary fluid
implementation, although the usual situation is to have only one.
The Model order for two $N$-velocity distributions $f$ and $g$ is:
\[
\underbrace{
\boxed{f_0\vphantom{f_{N-1}}}    \boxed{f_1\vphantom{f_{N-1}}}
\boxed{\ldots\vphantom{f_{N-1}}} \boxed{f_{N-1}} \,
\boxed{g_0\vphantom{f_{N-1}}}        \boxed{g_1\vphantom{f_{N-1}}}
\boxed{\ldots\vphantom{f_{N-1}}} \boxed{g_{N-1}\vphantom{f_{N-1}}}}_{r_0}
\ldots
\underbrace{
\boxed{f_0\vphantom{f_{N-1}}}    \boxed{f_1\vphantom{f_{N-1}}}
\boxed{\ldots\vphantom{f_{N-1}}} \boxed{f_{N-1}} \,
\boxed{g_0\vphantom{f_{N-1}}}        \boxed{g_1\vphantom{f_{N-1}}}
\boxed{\ldots\vphantom{f_{N-1}}} \boxed{g_{N-1}\vphantom{f_{N-1}}}}_{r_{n_s=1}}
\]

\textbf{ADDR\_MODEL\_R}: The Model R order is:

@@

The Model R picture, which targets coalescene, can be viewed as
naturally supporting SIMD vectorisation by virtue of allowing
contiguous access to individual quantities as a function of lattice
index. (If SIMD vectorisation is wanted at all on GPU architecture,
it can be as a means to releive register pressure.) Model R therefore
remains unaltered and we concentrate on the Model picture.

@@

\textbf{VADDR\_MODEL}: The structure of short vectors is based on the
SIMD vector length which we here denote $V$:
\[
\underbrace{
\overbrace{
\begin{array}{cc} \boxed{v_0} \\  _{r_0} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{v_0} \\  _{r_1} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{\ldots\vphantom{v_0}} \\ \phantom{r_2}\end{array} \mkern-20mu
\begin{array}{cc} \boxed{v_0} \\  _{r_{V-1}} \end{array}
}^{component_0}
\overbrace{
\begin{array}{cc} \boxed{v_1} \\  _{r_0} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{v_1} \\  _{r_1} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{\ldots\vphantom{v_0}} \\ \phantom{r_2}\end{array} \mkern-20mu
\begin{array}{cc} \boxed{v_1} \\  _{r_{V-1}} \end{array}
}^{component_1}
%\begin{array}{cc} \ldots \\ \phantom{r_2} \end{array}
\overbrace{
\begin{array}{cc} \boxed{v_{n-1}} \\  _{r_0} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{v_{n-1}} \\  _{r_1} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{\ldots\vphantom{v_{n-1}}} \\ \phantom{r_2}\end{array}
\mkern-17mu
\begin{array}{cc} \boxed{v_{n-1}} \\  _{r_{V-1}} \end{array}
}^{component_{n-1}}}_{\mathrm{SIMD\ BLOCK}}
\quad\ldots
\]
Subsequent SIMD blocks involve lattice sites $r_V \ldots r_{2V-1}$,
$r_{2V} \ldots r_{3V-1}$, and so on. If the SIMD vector length is unity,
this collapses to the standard Model picture shown in
Section~\ref{subsection:addressing-model-rank1}.

@@

@@

      index = coords_index(ic, jc, kc);
      /* Perform operation per lattice site ... */
    }
  }
}
\end{lstlisting}
where final array indexing is based on the coordinate index and is
specific to the object at hand. To allow transparent switching
between Model and Model R versions, the indexing must be abstracted.
As a concrete example, the following considers a rank 1 3-vector:
\begin{lstlisting}
for (ic = 1; ic <= nlocal[X]; ic++) {
  for (jc = 1; jc <= nlocal[Y]; jc++) {
    for (kc = 1; kc <= nlocal[Z]; kc++) {

@@

An equivalent vectorised version may be constructed as follows:
\begin{lstlisting}
for (n = 0; n < nsites; n += SIMDVL) {
  index = coords_indexv(n, nextra);       /* Proposal */
  for (ia = 0; ia < 3; ia++) {
    for (iv = 0; iv < SIMDVL; iv++) {
      iref = vaddr_rank1(nsites, 3, index, ia, iv);
      array[iref] = ...;
    }
  }
}
\end{lstlisting}
Note that a different function \texttt{vaddr\_rank1()} with an additional
argument which is the index of the vector loop is used. Note also that
vectorisaed versions must take into account the extent that the kernel
extends into the halo region (here \texttt{nextra}), which involves
logic to avoid lattice sites which are not required (not shown in this
example).

Note that, in practice, 
the loop over lattice sites is further abstracted in the code to
accommodate thread-level parallelism via targetDP interface.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  develop.tex
%
%  Further information for delvelopers
%
%  Edinburgh Soft Matter and Statistical Physics Group and
%  Edinburgh Parallel Computing Centre
%
%  (c) 2014-2017 The University of Edinburgh
%
%  Contributing authors:
%  Kevin Stratford (kevin@epcc.ed.ac.uk)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The code allows for different order of storage associated with 3-dimensional
fields (scalars, vectors, and so on). These storage schemes are usually
referred to as array of structures (AOS), structure of arrays (SOA), and
for the blocked version array of structures of (short) arrays (AOSOA).
For historical reasons these are
also sometimes referred to as `Model' and `Model~R' options, which correspond
to array-of-structures and structure-of-arrays layout, respectively. Early
versions of the code for CPU were written to favour summation of LB
distributions on a per lattice site basis in operations such as
$\rho(\mathbf{r}) = \sum_i f_i(\mathbf{r})$. This is array-of-structures,
where the $f_i$ are stored contiguously per site. Introduction of GPU
versions, where memory coalescing favours the opposite memory order,
were then referred to as `Model~R', the `R' standing for `reverse'.

@@

\textbf{ADDR\_AOS}: The array-of-structures order for an
$n-$vector field
with components $ v_\alpha = (v_0, v_1, \ldots, v_{n-1})$ is, schematically:
\[
\underbrace{ \boxed{v_0\vphantom{v_{n-1}}} \boxed{v_1\vphantom{v_{n-1}}}
\boxed{\ldots\vphantom{v_{n-1}}} \boxed{v_{n-1}} }_{r_0}
\underbrace{ \boxed{v_0\vphantom{v_{n-1}}} \boxed{v_1\vphantom{v_{n-1}}}
\boxed{\ldots\vphantom{v_{n-1}}} \boxed{v_{n-1}} }_{r_1}
\underbrace{ \boxed{v_0\vphantom{v_{n-1}}} \boxed{v_1\vphantom{v_{n-1}}}
\boxed{\ldots\vphantom{v_{n-1}}} \boxed{v_{n-1}} }_{r_2} \ldots
\underbrace{ \boxed{v_0\vphantom{v_{n-1}}} \boxed{v_1\vphantom{v_{n-1}}}
\boxed{\ldots\vphantom{v_{n-1}}} \boxed{v_{n-1}} }_{r_{n_s-1}}
\]

\textbf{ADDR\_SOA}: The structure-of-arrays version is:
\[
\overbrace{
\begin{array}{cc} \boxed{v_0} \\  _{r_0} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{v_0} \\  _{r_1} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{\ldots\vphantom{v_0}} \\ \phantom{r_2}\end{array} \mkern-22mu
\begin{array}{cc} \boxed{v_0} \\  _{r_{n_s-1}} \end{array}
}^{component_0}
\overbrace{
\begin{array}{cc} \boxed{v_1} \\  _{r_0} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{v_1} \\  _{r_1} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{\ldots\vphantom{v_0}} \\ \phantom{r_2}\end{array} \mkern-22mu
\begin{array}{cc} \boxed{v_1} \\  _{r_{n_s-1}} \end{array}
}^{component_1}
\begin{array}{cc} \ldots \\ \phantom{r_2} \end{array}
\overbrace{
\begin{array}{cc} \boxed{v_{n-1}} \\  _{r_0} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{v_{n-1}} \\  _{r_1} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{\ldots\vphantom{v_{n-1}}} \\ \phantom{r_2}\end{array} \mkern-17mu
\begin{array}{cc} \boxed{v_{n-1}} \\  _{r_{n_s-1}} \end{array}
}^{component_{n-1}}
\]

@@

\textbf{ADDR\_AOS}:
A general rank 2 tensor $t_{\alpha\beta}$ with components $(t_{0,0},
\ldots, t_{m-1,n-1})$ is stored as:
\begin{gather*}
\underbrace{
\boxed{t_{0,0}\vphantom{t_{m-1,n-1}}} \boxed{t_{1,0}\vphantom{t_{m-1,n-1}}}
\boxed{\ldots\vphantom{t_{m-1,n-1}}} \boxed{t_{m-1,0}\vphantom{t_{m-1,n-1}}}
\boxed{t_{0,1}\vphantom{t_{m-1,n-1}}} \boxed{t_{1,1}\vphantom{t_{m-1,n-1}}}
\boxed{\ldots\vphantom{t_{m-1,n-1}}} \boxed{t_{m-1,1}\vphantom{t_{m-1,n-1}}}
\ldots
\boxed{t_{0,n-1}\vphantom{t_{m-1,n-1}}}\boxed{t_{1,n-1} \vphantom{t_{m-1,n-1}}}
\boxed{\ldots\vphantom{t_{m-1,n-1}}} \boxed{t_{m-1,n-1}\vphantom{t_{m-1,n-1}}}
}_{r_0}
\ldots
\\
\ldots
\underbrace{
\boxed{t_{0,0}\vphantom{t_{m-1,n-1}}} \boxed{t_{1,0}\vphantom{t_{m-1,n-1}}}
\boxed{\ldots\vphantom{t_{m-1,n-1}}} \boxed{t_{m-1,0}\vphantom{t_{m-1,n-1}}}
\boxed{t_{0,1}\vphantom{t_{m-1,n-1}}} \boxed{t_{1,1}\vphantom{t_{m-1,n-1}}}
\boxed{\ldots\vphantom{t_{m-1,n-1}}} \boxed{t_{m-1,1}\vphantom{t_{m-1,n-1}}}
\ldots
\boxed{t_{0,n-1}\vphantom{t_{m-1,n-1}}}\boxed{t_{1,n-1} \vphantom{t_{m-1,n-1}}}
\boxed{\ldots\vphantom{t_{m-1,n-1}}} \boxed{t_{m-1,n-1}\vphantom{t_{m-1,n-1}}}
}_{r_{n_s-1}}
\end{gather*}
Dyadic tensors, for example the gradient of a vector field
$\partial_\alpha v_\beta$ in three dimensions, are stored in corresponding
fashion with $m=3$ and
$\partial_\alpha = (\partial_x, \partial_y, \partial_z)$.

\textbf{ADDR\_SOA}: The structure-of-arrays version is
\[
\overbrace{
\begin{array}{cc} \boxed{t_{0,0}} \\  _{r_0} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{t_{0,0}} \\  _{r_1} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{\ldots\vphantom{t_{0,0}}} \\ \phantom{r_2}\end{array} \mkern-18mu
\begin{array}{cc} \boxed{t_{0,0}} \\  _{r_{n_s-1}} \end{array}
}^{component_{0,0}}
\overbrace{
\begin{array}{cc} \boxed{t_{1,0}} \\  _{r_0} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{t_{1,0}} \\  _{r_1} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{\ldots\vphantom{t_{0,0}}} \\ \phantom{r_2}\end{array} \mkern-18mu
\begin{array}{cc} \boxed{t_{1,0}} \\  _{r_{n_s-1}} \end{array}
}^{component_{1,0}}
\begin{array}{cc} \ldots\vphantom{t_{n-1}} \\ \phantom{r_2}\end{array}
\overbrace{
\begin{array}{cc} \boxed{t_{m-1,n-1}} \\  _{r_0} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{t_{m-1,n-1}} \\  _{r_1} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{\ldots\vphantom{t_{0,0}}} \\ \phantom{r_2}\end{array} \mkern-17mu
\begin{array}{cc} \boxed{t_{m-1,n-1}} \\  _{r_{n_s-1}} \end{array}
}^{component_{m-1,n-1}}
\]

@@

The general rank 3 object $t_{\alpha\beta\gamma}$ with components
$(t_{0,0,0}, \ldots, t_{m-1,n-1,p-1})$ is stored in a manner which
generalises from the above, i.e., with the rightmost index running
fastest. Diagrams are omitted, but AOS and SOA storage patterns
follow the same form as seen above.

@@

\textbf{ADDR\_AOS}: For the LB distributions, up to two distinct
distributions can be accommodated to allow for a binary fluid
implementation, although the usual situation is to have only one.
The AOS order for two $N$-velocity distributions $f$ and $g$ is:
\[
\underbrace{
\boxed{f_0\vphantom{f_{N-1}}}    \boxed{f_1\vphantom{f_{N-1}}}
\boxed{\ldots\vphantom{f_{N-1}}} \boxed{f_{N-1}} \,
\boxed{g_0\vphantom{f_{N-1}}}        \boxed{g_1\vphantom{f_{N-1}}}
\boxed{\ldots\vphantom{f_{N-1}}} \boxed{g_{N-1}\vphantom{f_{N-1}}}}_{r_0}
\ldots
\underbrace{
\boxed{f_0\vphantom{f_{N-1}}}    \boxed{f_1\vphantom{f_{N-1}}}
\boxed{\ldots\vphantom{f_{N-1}}} \boxed{f_{N-1}} \,
\boxed{g_0\vphantom{f_{N-1}}}        \boxed{g_1\vphantom{f_{N-1}}}
\boxed{\ldots\vphantom{f_{N-1}}} \boxed{g_{N-1}\vphantom{f_{N-1}}}}_{r_{n_s=1}}
\]

\textbf{ADDR\_SOA}: The strucuture-of-arrays order is:

@@

The SOA picture, which targets coalescence, can be viewed as
naturally supporting SIMD vectorisation by virtue of allowing
contiguous access to individual quantities as a function of lattice
index. (If SIMD vectorisation is wanted at all on GPU architecture,
it can be as a means to relieve register pressure.) SOA therefore
remains unaltered and we concentrate on the AOS picture.

@@

\textbf{ADDR\_AOSOA}: The structure of short vectors is based on the
SIMD vector length which we here denote $V$:
\[
\underbrace{
\overbrace{
\begin{array}{cc} \boxed{v_0} \\  _{r_0} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{v_0} \\  _{r_1} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{\ldots\vphantom{v_0}} \\ \phantom{r_2}\end{array} \mkern-20mu
\begin{array}{cc} \boxed{v_0} \\  _{r_{V-1}} \end{array}
}^{component_0}
\overbrace{
\begin{array}{cc} \boxed{v_1} \\  _{r_0} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{v_1} \\  _{r_1} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{\ldots\vphantom{v_0}} \\ \phantom{r_2}\end{array} \mkern-20mu
\begin{array}{cc} \boxed{v_1} \\  _{r_{V-1}} \end{array}
}^{component_1}
%\begin{array}{cc} \ldots \\ \phantom{r_2} \end{array}
\overbrace{
\begin{array}{cc} \boxed{v_{n-1}} \\  _{r_0} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{v_{n-1}} \\  _{r_1} \end{array} \mkern-17mu
\begin{array}{cc} \boxed{\ldots\vphantom{v_{n-1}}} \\ \phantom{r_2}\end{array}
\mkern-17mu
\begin{array}{cc} \boxed{v_{n-1}} \\  _{r_{V-1}} \end{array}
}^{component_{n-1}}}_{\mathrm{SIMD\ BLOCK}}
\quad\ldots
\]
Subsequent SIMD blocks involve lattice sites $r_V \ldots r_{2V-1}$,
$r_{2V} \ldots r_{3V-1}$, and so on. If the SIMD vector length is unity,
this collapses to the standard Model picture shown in
Section~\ref{subsection:addressing-model-rank1}.

@@

\subsubsection{Compilation}

A number of options exist to control at compilation time the choice
of the address model. If no options are specified the default
addressing model is AOS. Two additional C preprocessor directives
exist:

\texttt{-DADDR\_SOA} selects structure-of-arrays order (e.g., for
coallescing on GPU platforms).

\texttt{-DADDR\_AOSOA} block SOA pattern (usually combined with
a non-unit vector length).

For development purposes, the address functions allow the array indices
to be checked against the correct range. For production runs, the
relevant functions are replaced by macros if \texttt{-DNDEBUG} is
specified.

The targetDP vector length is specified via, e.g.,  \texttt{-DVVL=4}.

\subsubsection{Kernels}

@@

      index = coords_index(ic, jc, kc);
      /* Perform operation per lattice site ... */
    }
  }
}
\end{lstlisting}
where final array indexing is based on the coordinate index and is
specific to the object at hand. To allow transparent switching
between different addressing models, the indexing must be abstracted.
The abstraction is provided in \texttt{memory.h}, where functions
(or function-like macros) are provided to make the appropriate
address computation.
As a concrete example, the following considers a rank 1 3-vector:
\begin{lstlisting}
for (ic = 1; ic <= nlocal[X]; ic++) {
  for (jc = 1; jc <= nlocal[Y]; jc++) {
    for (kc = 1; kc <= nlocal[Z]; kc++) {

@@

This will usually be carried out in computational kernels, where the
\texttt{kernel.h} interface can be used to help to manage the
one-dimensional loop structure required the the targetDP threading
model. In this way, the kernel launch stage will specify the limits
of the kernel in terms of local $x$-coordinates, $y$-coordinates, and
$z$-coordinates in the usual way. The kernel itself is then written
in terms of a one-dimensions kernel index.

In the following schematic, the launch stage defines a kernel context
to be passed to the kernel along with the other actual arguments:
\begin{lstlisting}
  kernel_info_t limits;
  kernel_ctxt_t * ctxt;

  limits.xmin = 1; limits.xmax = nlocal[X];
  limits.ymin = 1; limits.ymax = nlocal[Y];
  limits.zmin = 1; limits,zmax = nlocal[Z];

  kernel_ctxt_create(NSIMDVL, limits, &ctxt);

  __host_launch(kernel_function, nblk, ntpb, ktx, ...);
\end{lstlisting}

An suitable  vectorised kernel may be constructed as follows with the
appropriate base index for the loop computed via the utility
\texttt{kernel\_baseindex()}:
\begin{lstlisting}
for (kindex = 0; kindex < kiterations; kindex += SIMDVL) {
  index = kernel_baseindex(ctxt, kindex);
  for (ia = 0; ia < 3; ia++) {
    for (iv = 0; iv < SIMDVL; iv++) {
      iref = addr_rank1(nsites, 3, index + iv, ia);
      array[iref] = ...;
    }
  }
}
\end{lstlisting}
Note that the same function \texttt{addr\_rank1()} with the 
argument which is the index of the vector loop is used. Note also that
vectorised versions must take into account the extent that the kernel
extends into the halo region, which involves
logic to avoid lattice sites which are not required (this type of
masking operation is not shown in this
example).
